---
title: "01 Introduction to modeling"
author: "Edneide Ramalho"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    html_document:
      highlight: textmate
      logo: logo.png
      theme: cerulean
      number_sections: yes
      toc: yes
      toc_float:
        collapsed: yes
        smooth_scroll: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exercises

## **Exploratory visualization of age**

Let's perform an exploratory data analysis (EDA) of the numerical explanatory variable `age`. You should always perform an exploratory analysis of your variables before any formal modeling. This will give you a sense of your variable's distributions, any outliers, and any patterns that might be useful when constructing your eventual model.

-   Using the `ggplot2` package, create a histogram of `age` with bins in 5 year increments.

-   Label the `x` axis with `"age"` and the `y` axis with `"count"`.

```{r}
# Load packages
library(moderndive)
library(ggplot2)

# Plot the histogram
ggplot(evals, aes(x = age)) +
  geom_histogram(binwidth = 5) +
  labs(x = "age", y = "count")
```

## **Numerical summaries of age**

Let's continue our exploratory data analysis of the numerical explanatory variable `age` by computing **summary statistics**. Summary statistics take many values and summarize them with a single value. Let's compute three such values using `dplyr` data wrangling: mean (AKA the average), the median (the middle value), and the standard deviation (a measure of spread/variation).

Calculate the mean, median, and standard deviation of `age`.

```{r}
# Load packages
library(moderndive)
library(dplyr)

# Compute summary stats
evals %>%
  summarise(mean_age = mean(age),
            median_age = median(age),
            sd_age = sd(age))
```

# Background on modeling for prediction

**Question**:

> Can we predict the sale price of houses based on their features?

**Variables:**

-   $y$: House sale price

-   $\vec{x}$: Features like `sqft_living`, `condintion`, `bedrooms`, `yr_built`, `waterfront`

```{r}
library(dplyr)
library(moderndive)
glimpse(house_prices)
```

Steps to follow before modeling:

-   Looking at the data

-   Visualization

-   Summary Statistics

## EDA

```{r}
library(ggplot2)
ggplot(house_prices, aes(x = price)) +
  geom_histogram() +
  labs(x = "house price", y = "count")
```

-   The majority of the houses is less than 2 million dollars

-   There are few houses with prices close to 8 million dollars

-   The price distribution is right-skewed (it shows a long right tail )

## Log10 Transformation

```{r}
house_prices <- house_prices %>% 
  mutate(log10_price = log10(price)) 
```

```{r}
ggplot(house_prices, aes(x = log10_price)) +
  geom_histogram() +
  labs(x = "log10 house price", y = "count")
```

-   With the log transformation, price distribution is less skewed and more symmetrical.

## Exercises

# **Exploratory visualization of house size**

Let's create an exploratory visualization of the predictor variable reflecting the size of houses: `sqft_living` the square footage of living space where 1 sq.foot ≈ 0.1 sq.meter.

After plotting the histogram, what can you say about the distribution of the variable `sqft_living`?

##### 

-   Create a histogram of `sqft_living`.

-   Label the `x` axis with `"Size (sq.feet)"` and the `y` axis with `"count"`.

```{r}
# Load packages
library(moderndive)
library(ggplot2)

# Plot the histogram
ggplot(house_prices, aes(x = sqft_living)) +
  geom_histogram() +
  labs(x = "Size (sq.feet)", y = "count")
```
